---
title: "Sta 523 - Final Exam - Fall 2015"
author: "Lin Xiao"
output: html_document
runtime: shiny
---


First check and install all the packages we need for the exam

```{r}
# Check for libraries and install #
listOfPackages <- c("shiny", "dplyr", "ggmap", "jsonlite", "maps", "magrittr", 
"MazamaSpatialUtils", "stringr", "maptools", "testthat")
NewPackages <- listOfPackages[!(listOfPackages %in% installed.packages()[,"Package"])]
if(length(NewPackages)>0) {install.packages(NewPackages,repos="http://cran.rstudio.com/")}

suppressMessages(suppressWarnings(library(shiny)))
suppressMessages(suppressWarnings(library(testthat)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggmap)))
suppressMessages(suppressWarnings(library(jsonlite)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(maps)))
suppressMessages(suppressWarnings(library(maptools)))
suppressMessages(suppressWarnings(library(MazamaSpatialUtils)))
suppressMessages(suppressWarnings(library(stringr)))
```


1. Sparse Matrices

Task 1 - Validation

Write a function that will validate if the supplied object is a valid sparse matrix (i.e. meets all of the requirements above).

Here I use `expect` series of function within `testthat` pacakge to help me validate the sparse matrix, if all the tests are passed, a True will be returned, else error will be announced.

```{r}
# vsm stands for valid sparse matrix
validation <- function(vsm){
  # Structure is a list
  expect_is(vsm, "list")
  
  # Three elements
  expect_true(length(vsm) == 3)
  
  # One of the elements is "nrow"
  expect_true("nrow" %in% names(vsm))
  
  # Is an integer
  expect_true(vsm$nrow %% 1 == 0)
  
  # Integer > 0
  expect_true(vsm$nrow > 0)
  
  # One of the elements is "ncol"
  expect_true("ncol" %in% names(vsm))
  
  # Is an integer
  expect_true(vsm$ncol %% 1 == 0)
  
  # Integer > 0
  expect_true(vsm$ncol > 0)
  
  # Leave the data frame alone for examination
  vsm$nrow <- NULL
  vsm$ncol <- NULL
  
  # The last element is a data frame
  expect_is(vsm[[1]], "data.frame")
  
  # Three columns: ‘row’, ‘col’, ‘value’ 
  expect_true(setequal(colnames(vsm[[1]]), c("row", "col", "value")))
  
  # ‘row’ and ‘col’ should be integers
  expect_true(all(vsm[[1]][,1] %% 1 == 0) == T)
  expect_true(all(vsm[[1]][,2] %% 1 == 0) == T)
  
  # ‘value’ should be numeric
  expect_true(all(is.numeric(vsm[[1]][,3])) == T)
  
  # Ordered by row value and then col value
  expect_identical(vsm[[1]], arrange(vsm[[1]], row, col))
  return(T)
}
```

Task 2 - Transform sparse matrix

Write a function that accepts a sparse matrix as its argument and returns a dense matrix.


To transform a sparse matrix, that is, a list, into dense matrix, we first construct a matrix and setting the number of columns and rows to be the ones in the list. Then we just simply loop through all the rows in the data frame of the list, and label the value according to the giving 'row' and 'col'.

```{r}
transform_sparse_matrix <- function(vsm){
  # if it's valid, elso error will be returned
  if(validation(vsm) == T){
    nrow <- vsm$nrow
    ncol <- vsm$ncol
    
    # Leave the data frame alone for later use
    vsm$nrow <- NULL
    vsm$ncol <- NULL
    sparse_mat <- vsm[[1]]
    dense_mat <- matrix(0, nrow = nrow, ncol = ncol)
    for(i in 1:nrow(sparse_mat)){
      dense_mat[sparse_mat[i,1],sparse_mat[i,2]] <- sparse_mat[i,3]
    }
    return(dense_mat)
  }
}
```

An error will be returned if the sparse matrix is not valid.

Task 3 - Transform dense matrix

Write a function that accepts a dense matrix as its argument and returns a sparse matrix. Note that if we transform from dense -> sparse -> dense or sparse -> dense -> sparse that starting object should be identical to ending object.

Fisrt we get the number of rows and columns and store them in the list. Then we use use which function or just subset the indices and values. At last we need to reorder the rows and columns to make sure it's valid.

```{r}
transform_dense_matrix <- function(dm){
  expect_true(is.matrix(dm))
  # Return the indices with their non-zero values
  sparse_mat <- data.frame(which(dm != 0, arr.ind = T), value = dm[dm != 0])
  
  # Order by row and then col
  sparse_mat <- arrange(sparse_mat, row, col)
  valid_sparse_matrix <- list(nrow = nrow(dm), ncol = ncol(dm), sparse_mat)
  
  # Test to check if they are equal, if not, error will be returned
  expect_equal(dm, transform_sparse_matrix(valid_sparse_matrix))
  return(valid_sparse_matrix)
}
```

Task 4 - Matrix Multiplication

Write a function that will accept two matrices as its arguments (either sparse or dense) and calculates the matrix product of the two input matrices. If multiplying two sparse matrices the returned matrix should also be sparse, you are also not allowed to convert these matrices into a dense form at any point. Any multiplication operation involving a dense matrix should return a dense matrix. Be sure to also check that the provided matrices are conforming.

If two dense matrices are the inputs, we can do matrix multiplication easily. Else if two sparse matrices are the inputs, since we are not allowed to transform them into dense matrices and do the simple calculation, we can loop through the data frames and connect them to calculate the dense matrix we want and later transform it back to the sparse matrix. For details, we loop through the rows of the left matrix. For each row in it, we search for the value in the 'col' column in the 'row' column of the data frame in the right sparse matrix. Once connected we multiply the values together and assign it to the new dense matrix, which is defined with the nrow from the left and the ncol from the right, and the corresponding indices of this multiplied value are the value in the 'row' of the left sparse matrix, and the value in the 'col' of the right sparse matrix. We add all the values into the dense matrix through looping and obtain the matrix we need. The inputs can also be a sparse matrix and a dense matrix, we can transform the dense matrix into a sparse matrix and repeat the same thing as above.

```{r}
multiplication <- function(m1, m2){
  
  # If the inputs are both sparse "matrix"
  if(is.list(m1) & is.list(m2)){
    
    # Check validity first
    stopifnot(validation(m1) & validation(m2)) # Return error
    if(m1$ncol != m2$nrow) return(F) # Return False
    res_mat <- matrix(0, nrow = m1$nrow, ncol = m2$ncol)
    
    # Leave the data frame for later use
    m1$nrow <- NULL
    m1$ncol <- NULL
    m2$nrow <- NULL
    m2$ncol <- NULL
    rec_mat1 <- m1[[1]]
    rec_mat2 <- m2[[1]]
    
    # Main step, as described above
    for(i in 1:nrow(rec_mat1)){
      for(j in which(rec_mat2[,1] == rec_mat1[i,2])){
        res_mat[rec_mat1[i,1],rec_mat2[j,2]] <- 
          res_mat[rec_mat1[i,1],rec_mat2[j,2]] + 
          rec_mat1[i,3]*rec_mat2[j,3]
      }
    }
    
    # Transform back to sparse "matrix"
    result_sparse_mat <- transform_dense_matrix(res_mat)
    return(result_sparse_mat)
  }
  
  # If inputs are a sparse and a dense matrix
  else if(is.list(m1) & is.matrix(m2)){
    multiplication(m1, transform_dense_matrix(m2))
  }else if(is.list(m2) & is.matrix(m1)){
    multiplication(m2, transform_dense_matrix(m1))
  }
  
  # If both are dense matrix
  else if(is.matrix(m1) & is.matrix(m2)){
    expect_equal(ncol(m1), nrow(m2))
    return(m1 %*% m2)
  }
}
```

Task 5 - testthat

Write at least 5 tests for each of the above functions using the testthat package. Note that you shouldn’t just test for correct behavior - check that errors are handled correctly.

```{r}
# Write some samples for testing
tdmatrix <- matrix(c(1,0,0,3,0,2,0,1,3,8,0,0), ncol = 3)
tdmatrix1 <- matrix(c(1,0,3,3,0,2,0,1,3,8,0,0), ncol = 3)
tdmatrix2 <- matrix(c(1,0,9,3,0,2,0,1,3,8,0,0), ncol = 3)
mat <- data.frame(matrix(c(1,1,1,1,3,3,2,2,2,2,3,8,4,1,3,4,2,1), ncol = 3, byrow = T))
colnames(mat) <- c("row", "col", "value")
tsmatrix1 <- list(ncol = ncol(tdmatrix), nrow = nrow(tdmatrix), mat)
tsmatrix2 <- list(nrow = nrow(tdmatrix), ncol = ncol(tdmatrix), mat)
tsmatrix3 <- list(nroww = nrow(tdmatrix), ncol = ncol(tdmatrix), mat)
tsmatrix4 <- list(nrow = nrow(tdmatrix)+1, ncol = ncol(tdmatrix), mat)
tsmatrix5 <- list(nrow = 5, ncol = 5,data.frame( mat = matrix(c(1,1,1), ncol = 3, nrow = 1, byrow = T)))
tsmatrix5 <- list(nrow = 5, ncol = 5, value = data.frame( mat = matrix(c(1,1,1), ncol = 3, nrow = 1, byrow = T)))



# Test for validation function
expect_error(validation(tdmatrix))
expect_true(validation(tsmatrix1))
expect_true(validation(tsmatrix2))
expect_error(validation(tsmatrix3))
expect_true(validation(tsmatrix4))

# Test for transform_sparse_matrix function
expect_identical(transform_sparse_matrix(tsmatrix1), tdmatrix)
expect_identical(transform_sparse_matrix(tsmatrix2), tdmatrix)
expect_error(transform_sparse_matrix(tsmatrix3))
expect_false(any(transform_sparse_matrix(tsmatrix1) == (tdmatrix-1)))
expect_error(transform_sparse_matrix(tsmatrix5))


# Test for transform_dense_matrix function
expect_equal(transform_dense_matrix(tdmatrix)[[3]],tsmatrix1[[3]])
expect_equal(transform_dense_matrix(tdmatrix)[[3]],tsmatrix2[[3]])
expect_equal(transform_dense_matrix(tdmatrix)$nrow,tsmatrix1$nrow)
expect_error(transform_dense_matrix(tsmatrix))
expect_is(transform_dense_matrix(tdmatrix), "list")

# Test for multiplication function
expect_false(multiplication(tsmatrix1, tsmatrix1))
expect_error(multiplication(tdmatrix, tdmatrix))
expect_equivalent(multiplication(t(tdmatrix), tdmatrix), transform_sparse_matrix(multiplication(transform_dense_matrix(t(tdmatrix)), transform_dense_matrix(tdmatrix))))
expect_equivalent(multiplication(t(tdmatrix1), tdmatrix1), transform_sparse_matrix(multiplication(transform_dense_matrix(t(tdmatrix1)), transform_dense_matrix(tdmatrix1))))
expect_equivalent(multiplication(t(tdmatrix2), tdmatrix2), transform_sparse_matrix(multiplication(transform_dense_matrix(t(tdmatrix2)), transform_dense_matrix(tdmatrix2))))
```


2. Shiny Weather

```{r, echo=FALSE}

################################ 
### Shiny App
################################ 

key = "72dd839e601dd06f1e48e16ec40604c9"

# Factors for the icon column
weather = c("clear-day", "clear-night", "rain", "snow", "sleet", "wind", 
            "fog", "cloudy", "partly-cloudy-day", "partly-cloudy-night")

# 76 Major cities
city <- c("Amsterdam", "Ankara", "Athens", "Atlantic City", "Baltimore",
          "Bangkok", "Beijing", "Berlin", "Berne",
          "Brussels", "Budapest", "Buenos Aires", "Cairo", 
          "Canberra", "Cape Town", "Chicago", "Cologne",
          "Copenhagen", "Damascus", "Delhi", "Dubai", 
          "Dublin", "Florence", "Genève", "Ha Noi", "Havana", "Helsinki", 
          "Hong Kong", "Honolulu", "Istanbul", "Jakarta", "Jerusalem", 
          "Kansas City", "Kathmandu", "Kuala Lumpur", "Lisbon", 
          "London", "Los Angeles", "Luxembourg", 
          "Madrid", "Manila", "Melbourne", "Mexico City", "Milan", 
          "Montreal", "Moscow", "Mumbai", "Munich", "Nazareth", 
          "New York", "Nice", "Osaka", "Ottawa", "Oslo",
          "Paris", "Philadelphia", "Phnom Penh", "Prague", "Quito", 
          "Reykjavík", "Rio de Janeiro", "San Francisco", "Santiago", 
          "São Paulo", "Shanghai", "Singapore", "Stockholm", "Saint-Petersburg", 
          "Sydney", "Taipei", "Tokyo", "Toronto", "Venice", 
          "Vienna", "Washington", "Zürich")

# Attributes: Second quantity 
attributes <- c("summary", "icon", "precipIntensity", "precipProbability", 
                "apparentTemperature", "dewPoint", "humidity", "windSpeed", 
                "windBearing", "visibility", "cloudCover", "pressure", "ozone", "precipType")

```


Task 1 - Getting data from forecast.io

When selecting a city through Shiny, a corresponding web page will be scraped using fromJSON. Since we only need hourly data, we can use `fromJSON(url)$hourly` to get the list of data, and later a summary and a data frame of the list will be returned for later use. Note that the time format of the data is just a bunch of numbers indicating the seconds since 1970-01-01, we need to make a transformation on it.

Task 2 - Shiny Predictions


We have three plots to show, major plot with temperature on it, "other plot" which will always return the second quantity plot user has chosen, and the world map indicating the location of the city. "Other plot" is being used here because some columns have categorical data, we cannot plot them with the first plot. Otherwise, numeric lines will be provided in these two plots. 

For second quantities, icon is a factor with ten types of weather, preciType is not shown if there is 0 probability for precipitation, and for summary column the document doesn't give enough detail about it, so we should deal with them separately in the code. If user selects categorical data, the major plot will only have temperature on it and the selected data will be shown in the "other plot".

The longitude and the latitude, along with the summary for the current are scraped and being plotted on the screen.

```{r, echo=FALSE}
shinyApp(
  ui =   fluidPage(
    # Give the panel a title
    titlePanel(
       "Weather forecast"
    ),
    # Use fluidRow here for better design
    fluidRow(
      column(5, align = "center",
             h2("Major cities:"),
             selectInput("Cities", label = "Choose a city to display", choices = city),
             hr(),
             # Select the attributes for second quantity
             selectInput("Attributes", label = "Choose a second quantity you want to display",
                         choices = attributes, selected = "precipIntensity")),
      column(5,
             # Summary is from the original hourly data list
             h3("Summary"),
             h5(textOutput("summary")),
             # Use geocoding to get the longitude and latitude
             h3("Location"),
             h5(textOutput("location")),
             # Return the country
             h3("Country"),
             h5(textOutput("country"))
      )
    ),
    ### Show results in main panel
    br(),
    hr(),
    fluidRow(
      h2("Results:"),
      plotOutput("temperature_plot")
    ),
      
    fluidRow(
      plotOutput("other_plot")
    ),
    fluidRow(
      plotOutput("Worldmap")
    )
),

#################################################################
#  We have three plots above, major plot with temperature on it,
#  Other plot which will always return the second quantity plot
#  user has chosen, and the world map indicating the location of
#  the city. Other plot is being used here because some columns 
#  have categorical data, we cannot plot them with the first plot.
#  Otherwise, numeric lines will be provided in these two plots.
################################################################


  server = function(input, output, session){
    
    
    # Return a data list first
    datalist = reactive(
      { 
        loc = geocode(input$Cities, source = "google")
        url = paste0("https://api.forecast.io/forecast/", key, "/", loc[[2]], ",", loc[[1]])
        dat = fromJSON(url)$hourly
        dat
      }
    )
    
    # Extract the summary
    summary = reactive(
      {
        datalist()$summary
      }
    )
    
    # Extract the data frame
    data = reactive(
      {
        dat = datalist()$data
        dat$time = as.POSIXct(dat$time, origin = "1970-01-01")
        dat
      }
    )
    
    # Temperature plot
    output$temperature_plot = renderPlot(
      {
        
        
        # To make the time format shorter
        t <- data()$time %>% 
          str_replace("2015-12", "Dec") %>% 
          str_replace(":00:00", "") %>% 
          str_replace("-", "/")
        par(mar = c(8,10,4,10))
        
        # If the second quantity selected is continuous data, we
        # can plot it with temperature
        if(input$Attributes %in% c("precipIntensity", "precipProbability", 
                                   "apparentTemperature", "dewPoint", "humidity", 
                                   "windSpeed", "windBearing", "visibility", 
                                   "cloudCover", "pressure", "ozone")){
          plot(data()$time, data()$temperature, xaxt = "n", xlab = "", ylab = "temperature(F)", 
               main = paste0("temperature(F) and ", input$Attributes, " hourly predictions"),
               col = "red", type = "b")
          par(new = T)
          plot(data()$time, data()[,input$Attributes], ylab = " ", xlab = " ", axes = F, 
               col = "blue", type = "b")
          axis(1, labels = t, at = data()$time, las = 2)
          mtext("hour", side = 1, line = 6)
          mtext(input$Attributes, side = 4, line = 5) 
          axis(4, seq(range(data()[,input$Attributes])[1], 
                      range(data()[,input$Attributes])[2], length.out = 5), las = 1)
          legend("topright", c("temperature", input$Attributes),lty=c(1,1),
                 lwd=c(2.5,2.5), col=c("red", "blue"), cex=1.2)
        }
        # If not, we will stick to the temperature.
        else{
          plot(data()$time, data()$temperature, xaxt = "n", xlab = "", 
               ylab = "temperature(F)", main = "temperature(F) hourly predictions",
               col = "red", type = "b")
          axis(1, labels = t, at = data()$time, las = 2)
          mtext("hour", side=1, line = 6)
        }
      }
    )
    
    # Other plot: for second quantity
    output$other_plot = renderPlot(
      {
        # To make the time format shorter
        t <- data()$time %>% 
          str_replace("2015-12", "Dec") %>% 
          str_replace(":00:00", "") %>% 
          str_replace("-", "/")
        
        # If the second quantity selected is continuous data, plot the line
        if(input$Attributes %in% c("precipIntensity", "precipProbability", 
                                   "apparentTemperature", "dewPoint", "humidity", "windSpeed", 
                                   "windBearing", "visibility", "cloudCover", "pressure", "ozone")){
          par(mar = c(8,5,3,4))
          plot(data()$time, data()[,input$Attributes], xaxt = "n", xlab = "", 
               ylab = input$Attributes, main = paste0(input$Attributes, " hourly predictions"),
               col = "blue", type = "b")
          axis(1, labels = t, at = data()$time, las = 2)
          mtext("hour", side=1, line = 6)
        }
        # If it's icon, first factorize it, then return a scatterplot
        else if(input$Attributes == "icon"){
          par(mar = c(8,8,3,4))
          Response <- factor(data()$icon, levels = weather, labels = 1:10)
          plot(data()$time, Response, ylim = c(1,10), xaxt = "n", yaxt = "n",
               xlab = "", ylab = input$Attributes)
          axis(1, labels = t, at = data()$time, las = 2)
          mtext("hour", side=1, line = 6)
          axis(2, at = 1:10, labels = weather, las = 1)
        }
        # If it's preciptype, which is not in the data of every city,
        # fisrt give "NA" for the missing data
        # Then make a barplot
        else if(input$Attributes == "precipType"){
          pre <- data()[,input$Attributes] 
          pre[is.na(pre)] <- "NA"
          barplot(table(pre))
        }
        
        # If not return a barplot for summary
        else{
          barplot(table(data()[,input$Attributes]))
        }
      }
    )
    
    # Print the summary words
    output$summary = renderText({
      summary()
    })
    
    # Print the longitudes and latitudes
    output$location = renderText({
      loc = geocode(input$Cities, source = "google")
      paste0("Longitude is ", loc[[1]], " and latitude is ", loc[[2]])
    })
    
    # Print the country name
    output$country = renderText(
      {
        loc = geocode(input$Cities, source = "google")
        getCountryName(loc[[1]], loc[[2]])
      }
    )
    
    
    # Print the world map and use a red dot to indicate the location of the city
    output$Worldmap = renderPlot(
      {
        ll.visited <- geocode(input$Cities, source = "google")
        visit.x <- ll.visited$lon
        visit.y <- ll.visited$lat
        map("world", fill=TRUE, col="white", bg="grey", ylim=c(-60, 90), mar=c(0,0,0,0))
        points(visit.x,visit.y, col="red", pch = 8, cex = 2)
      }
    )
    
  }
)
```

Because of the html format, the plots are not as nice as they were supposed to be.


